Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-23 files/Chapter3/dynamicconvergence.tex
chapter 3

\subsection{Formulation of the Galerkin approximation}
Recall the spaces $V$, $W$ and $X$ from Section \ref{ssec:existence:VariationlApproach} where $V \subset W \subset X$.

@@

Before the theory of \cite{BV13} can be discussed, some preliminary work is
necessary. The structure of this section is as follows. First the Galerkin
approximation is for Problem GVar is derived. Then a equivalent system of
ordinary differential equations is derived using the Finite Element Method.
Finally, the convergence of the Galerkin approximation is discussed using the
work of the article \cite{BV13}.

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ in the finite dimensional space $S^h$.

\subsection{System of ordinary differential equations}
Problem GVar$^h$ is equivalent to a system of second order differential equations.

Consider the standard FEM matrices defined by
\begin{eqnarray*}
	K_{ij} & = & b(\phi_j, \phi_i)\\
	C_{ij} & = & a(\phi_j, \phi_i)\\
	M_{ij} & = & c(\phi_j, \phi_i)\\
	F_{i}(t) & = & b(f(t), \phi_i)
\end{eqnarray*}

@@

Recall that $u^h(t) = \sum_{k} \bar{u}_k(t)\phi_k$.

\subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that \label{sym:baru}
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}^h_0 \textrm{ and }  \bar{u}(1) = \bar{u}^h_1 \label{ODE}
\end{eqnarray}
The function $F$ is the interpolation of the function $f$ into the discrete space $S^h$.

@@

\newtheorem{DC_Prop3}[DC_Prop2]{Proposition}
\begin{DC_Prop3}
	Suppose $M, K, C, F,\bar{u}^h_0$ and $\bar{u}^h_1$ are defined as above. Then, the function $u_{h}$ is a solution of Problem GVar$^{h}$ if and only if the function $\bar{u}$ is a solution of Problem ODE.
\end{DC_Prop3}

Proposition 2 provides a link between the solution of Problem ODE and the
solution of Problem GVar$^h$.

The following theorem from \cite{BV13} proves that if $f$ is continuous, then problem Problem GVar$^h$ has a unique solution.

\newtheorem{DC_Thm2}{Theorem}
\begin{DC_Thm2} \label{DC_THM2}
	If $f \in C(J,X)$, then there exists a unique solution $u_h \in C^2(J)$ for Problem GVar$^h$ for each $u_0^h$ and $u_1^h$ in $S^h$. If $f = 0$ then $u_h \in C^2((-\infty, \infty))$.
\end{DC_Thm2}

Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length $\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for $k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \ t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.  A finite difference method is used to compute $u^h_k$ for each $k$.

@@

@@

From Theorem 2 of the semi-discrete problem an error estimate for
$e_k = [u(t_k)-u_h(t_k)]$ with respect to the norm of $W$ was obtained. The authors of \cite{BV13} then proceed to obtain an error estimate for $[u_h(t_k) - u^h_k]$.

Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $[u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$.

A stability result is needed.

\newtheorem*{DC_Lem3}{Lemma}
\begin{DC_Lem3}
	Stability
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq KT\tau
	\end{eqnarray*}
	where $K$ depends on $u$, $u_h$ and their derivatives.
\end{DC_Lem3}

Using this lemma, \cite{BV13} prove the finial error estimate.

>>>>

\subsection{Formulation of the Galerkin approximation}
Recall the spaces $V$, $W$ and $X$ from Section \ref{sec:existence:ExistenceAndUniqueness} where $V \subset W \subset X$.

@@

Before the theory of \cite{BV13} can be discussed, some preliminary work is
necessary. The structure of this section is as follows. First the Galerkin
approximation for Problem GVar is derived. Then an equivalent system of
ordinary differential equations is derived using the Finite Element Method.
Finally, the convergence of the Galerkin approximation is discussed using the
work of the article \cite{BV13}.

\textcolor{red}{*******NEED COME BACK TO THIS**********}

Let $S^h$ denote a finite dimensional subspace of $V$.

\textcolor{red}{*******NEED COME BACK TO THIS**********}

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J, S^h)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ in the finite dimensional space $S^h$.

\subsection{System of ordinary differential equations}
Problem GVar$^h$ is equivalent to a system of second order differential equations. Consider the standard FEM matrices defined by
\begin{eqnarray*}
	K_{ij} & = & b(\phi_j, \phi_i),\\
	C_{ij} & = & a(\phi_j, \phi_i),\\
	M_{ij} & = & c(\phi_j, \phi_i),\\
	F_{i}(t) & = & c(f(t), \phi_i),
\end{eqnarray*} where $\phi_i$ and $\phi_j$ are admissible basis functions.

@@

Recall that $u^h(t) = \sum_{k} u_k(t)\phi_k$ where $\bar{u}_k = \left( u_1(t), u_2(t), ...  , u_n(t)\right)$.

\subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that \label{sym:baru}
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}_0 \textrm{ and }  \bar{u}(1) = \bar{u}_1 \label{ODE}
\end{eqnarray}

@@

\newtheorem{DC_Prop3}[DC_Prop2]{Proposition}
\begin{DC_Prop3}
	Suppose $M, K, C, F,{u}_0$ and ${u}_1$ are defined as above. Then, the function $u_{h}$ is a solution of Problem GVar$^{h}$ if and only if the function $\bar{u}$ is a solution of Problem ODE.
\end{DC_Prop3}

Proposition 2 provides a link between the solution of Problem ODE and the
solution of Problem GVar$^h$. Theorem 1 below follows.

\newtheorem{DC_Thm2}{Theorem}
\begin{DC_Thm2} \label{DC_THM2}
	If $f \in C(J,X)$, then there exists a unique solution $u_h \in C^2(J, S^h)$ for Problem GVar$^h$ for each $u_0^h$ and $u_1^h$ in $S^h$. If $f = 0$ then $u_h \in C^2((-\infty, \infty))$.
\end{DC_Thm2}

It is now required to find an approximation for the solution of $\bar{u}$ of Problem ODE.

Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length $\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for $k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \ t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$. (Recall that $u_h(t) = \sum u_i(t)\phi_i$.) A finite difference method is used to compute $u^h_k$ for each $k$.

@@

\vspace*{40mm}

@@

From Theorem 2 of the semi-discrete problem an error estimate for
$e_k = [u(t_k)-u_h(t_k)]$ with respect to the norm of $W$ was obtained. The authors of \cite{BV13} then proceed to obtain an error estimate for $e_k = [u_h(t_k) - u^h_k]$.

Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $e_k = [u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$.

A stability result is derived in \cite{BV13}.

\newtheorem*{DC_Lem3}{Lemma}
\begin{DC_Lem3}
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq KT\tau
	\end{eqnarray*}
	where $K$ depends on $u$, $u_h$ and their derivatives.
\end{DC_Lem3}

Using this lemma, \cite{BV13} prove the error estimate.

\newtheorem*{DC_Thm5_1}{Theorem 2}
\begin{DC_Thm5_1}
	{Main Result}\\
	If $f\in C^{2}([0,T],X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & K\tau^2
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5_1}

### 2023-11-23 files/Chapter3/femconvergence.tex
chapter 3

The following eigenvalue is considered in \cite{SF73}.

\subsection{General eigenvalue problem}
\subsubsection*{Problem E}
Find a vector $u \in V$ and number $\lambda \in R$ such that $u \neq 0$ and
\begin{equation}
	b(u,v) = \lambda (u,v) \label{GEVP}
\end{equation} for each $v \in V$.

Consider the Galerkin approximation for \eqref{GEVP}:
\subsubsection*{Problem $E_h$}
Find $u^h \in S^h$ such that $u^h \neq 0$ and \[b(u^h, v) = \lambda^h(u^h,v) \textrm{ for all } v \in S^h.\]

@@

\subsection{Preliminary theory and notation}
In this section, the work in the textbook \cite[p.228-236]{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for greater clarity.

@@

@@

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_i \in V | i = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_i \in V | i = 1,2,...,m\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

@@

\textbf{Rayleigh quotient}
\begin{equation*}
	R(v) = \frac{b(v,v)}{(v,v)} \ \text { for } v \in V.
\end{equation*} \label{sym:Rayleigh}

@@

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_i \in V | i = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_i \in V | i = 1,2,...,m\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

\subsection{Estimating the eigenvalues.}
Let $E_j \in V$ denote the eigenspace spanned by the exact eigenvectors $\left\{u_1,u_2,...,u_j \right\}$ for $j = 1,2,...,m$.

>>>>

\subsection{General eigenvalue problem}

The following eigenvalue problem is considered in \cite{SF73}.

\subsubsection*{Problem E}
Find a vector $u \in V$ and number $\lambda \in R$ such that $u \neq 0$ and
\begin{equation}
	b(u,v) = \lambda (u,v) \label{GEVP}
\end{equation} for each $v \in V$.

Suppose $E^h$ is a finite dimensional subspace of $V$. Consider the Galerkin approximation for \eqref{GEVP}:
\subsubsection*{Problem $E^h$}
Find $u^h \in S^h$ such that $u^h \neq 0$ and \[b(u^h, v) = \lambda^h(u^h,v) \textrm{ for all } v \in S^h.\]

@@

\textcolor{red}{********** need to move this ******************}

\subsection{Preliminary theory and notation}
In this section, the work in the textbook \cite{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for greater clarity.

@@

\textcolor{red}{********** need to move this ******************}

\subsubsection*{Properties of eigenvalues}
It is well known that the eigenvalues of a eigenvalue problem such as \eqref{GEVP} form an infinite sequence of eigenvalues $(\lambda_i)$. Assume that the eigenvectors can be ordered in such a way that $\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...$ where $\lambda_i$ are the corresponding eigenvalues.

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_k \in V | k = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_k \in V | k = 1,2,...,m\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

Problem $E^h$ can be written as a matrix eigenvalue problem,
\begin{eqnarray*}
	\lambda^nM\bar{u}_n = K\bar{u}_n.
\end{eqnarray*} The pair $(\lambda^n, \bar{u}_n)$ correspond to the pair $(\lambda^h_k, u^h_k)$ which is the solution of Problem $E^h$.

In a finite-dimensional subspace, the ordering of vectors is preserved as it is in the original space. Denote the normalized eigenvectors in the space $S^h$ as $u^h_k$ with corresponding eigenvalues $\lambda^h_i$ for $k = 1,2,...,m$.

@@

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_k \in V | k = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_k \in V | k = 1,2,...,m\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

@@

\subsection{Estimating the eigenvalues.}
\textbf{Rayleigh quotient}
\begin{equation*}
	R(v) = \frac{b(v,v)}{(v,v)} \ \text { for } v \in V.
\end{equation*} \label{sym:Rayleigh}

@@

Let $E_j \in V$ denote the eigenspace spanned by the exact eigenvectors $\left\{u_1,u_2,...,u_j \right\}$ for $j = 1,2,...,m$.
